---
title: "Assignment 2A"
author: "Jose Vilardy, Caro Paipa, Jacky Wong, Justin Mah, Seeneyen Youshinee"
date: "2 September 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

There are two types of tables to consolidate for this research:

Table 1: Contains information as Average income, number of individuals, total income by the tax payers, number of individuals with health insurance, number of welfare recipients, total of welfare payment and year.

Table 2: Contains all the type of welfare payments made by the Government each year (Department of Social Services Payment Demographic Data)

The information for each table was structured by year as per the codes below:

# Table 1

### Table 1 Year 2013: Gathering Information for the 2013 Financial Year
```{r,  warning = FALSE}
library(readxl)
library(httr)
library(dplyr)
library(tidyr)

url2 <- "https://data.gov.au/dataset/e29ef9ca-0d1a-47ec-9e9b-14a79a941511/resource/fa6fda82-a70d-4f4c-bb00-203ffc4d20af/download/taxstats2013individual06selected-items-by-taxable-status-postcode.xlsx"
p2f <- tempfile()

GET(url2, write_disk(tf2 <- tempfile(fileext = ".xlsx")))

T2013 <- as.data.frame(read_excel(tf2, sheet = 3, col_names = TRUE, skip = 2))


T2013 <- T2013%>%select("State/Territory1", "Postcode", "Number of individuals\r\nno.", "Number of individuals with private health insurance no.","Taxable income or loss\r\nno.", "Taxable income or loss\r\n$", "Gross interest\r\n$", 
"Australian Government allowances and payments like newstart, youth allowance and austudy payment\r\n$", "Australian Government allowances and payments like newstart, youth allowance and austudy payment\r\nno.")

colnames(T2013) <- c("State", "Postcode", "Population", "NoPeopleHI","NoPeople", "Income", "Interest", "Allowances", "NoAllowances")

rep2013 <- rep("2013", length(T2013$State))

T2013 <- T2013%>%mutate(Year = rep2013)

```


### Table 1 Year 2014: Gathering Information for the 2014 Financial Year
```{r,  warning = FALSE}

url3 <- "https://data.gov.au/dataset/25e81c18-2083-4abe-81b6-0f530053c63f/resource/942c9a30-baf2-4432-9513-1395fbef7edf/download/taxstats2014individual06taxablestatuspostcode.xlsx"
p3f <- tempfile()

GET(url3, write_disk(tf3 <- tempfile(fileext = ".xlsx")))

T2014 <- as.data.frame(read_excel(tf3, sheet = 2, col_names = TRUE, skip = 2))




T2014 <- T2014%>%select("State/Territory1", "Postcode", "Number of individuals", "Number of individuals with private health insurance","Taxable income or loss3\r\nno.", "Taxable income or loss3\r\n$", "Gross interest\r\n$", 
                        "Australian government allowances and payments\r\n$", "Australian government allowances and payments\r\nno.")


colnames(T2014) <- c("State", "Postcode", "Population", "NoPeopleHI","NoPeople", "Income", "Interest", "Allowances", "NoAllowances")

T2014 <- T2014%>%group_by(State, Postcode)%>% 
  summarize(Population = sum(Population), NoPeopleHI = sum(NoPeopleHI), NoPeople = sum(NoPeople), Income = sum(Income), Interest = sum(Interest), Allowances = sum(Allowances), NoAllowances = sum(NoAllowances))


rep2014 <- rep("2014", length(25))


T2014 <- as.data.frame(T2014%>%mutate(Year = rep2014, Postcode = as.character(Postcode)))

```



### Table 1 Year 2015: Gathering Information for the 2015 Financial Year
```{r,  warning = FALSE}

url4 <- "https://data.gov.au/dataset/5c99cfed-254d-40a6-af1c-47412b7de6fe/resource/12443334-2273-4e0f-a2e1-c9d0c1d8213b/download/taxstats2015individual06taxablestatusstateterritorypostcode.xlsx"
p4f <- tempfile()

GET(url4, write_disk(tf4 <- tempfile(fileext = ".xlsx")))

T2015 <- as.data.frame(read_excel(tf4, sheet = 3, col_names = TRUE, skip = 2))



T2015 <- T2015%>%select("State/ Territory1", "Postcode", "Number of individuals\r\nno.", "People with private health insurance\r\nno.","Taxable income or loss3\r\nno.", "Taxable income or loss3\r\n$", "Gross interest\r\n$", 
                        "Australian government allowances and payments\r\n$", "Australian government allowances and payments\r\nno.")


colnames(T2015) <- c("State", "Postcode", "Population", "NoPeopleHI","NoPeople", "Income", "Interest", "Allowances", "NoAllowances")

rep2015 <- rep("2015", length(T2015$State))

T2015 <- T2015%>%mutate(Year = rep2015)
```


### Table 1 Year 2016: Gathering Information for the 2015 Financial Year
```{r,  warning = FALSE}

url5 <- "https://data.gov.au/dataset/d170213c-4391-4d10-ac24-b0c11768da3f/resource/8c86b6d5-0485-4387-b12b-ad26a9da2033/download/taxstats2016individual06taxablestatusstateterritorypostcodetaxableincome.xlsx"
p5f <- tempfile()

GET(url5, write_disk(tf5 <- tempfile(fileext = ".xlsx")))

T2016 <- as.data.frame(read_excel(tf5, sheet = 3, col_names = TRUE, skip = 2))



T2016 <- T2016%>%select("State/ Territory1", "Postcode", "Number of individuals\r\nno.", "People with private health insurance \r\nno.","Taxable income or loss3 \r\nno.", "Taxable income or loss3 \r\n$", "Gross interest \r\n$", 
                        "Australian government allowances and payments \r\n$", "Australian government allowances and payments \r\nno.")


colnames(T2016) <- c("State", "Postcode", "Population", "NoPeopleHI","NoPeople", "Income", "Interest", "Allowances", "NoAllowances")

rep2016 <- rep("2016", length(T2016$State))

T2016 <- T2016%>%mutate(Year = rep2016)

```


### Consolidating years 2013, 2014, 2015 and 2016 in Table 1

```{r,  warning = FALSE}

table1 <- rbind(T2013, T2014, T2015, T2016)

table1 <- table1%>%mutate(AverageIncome = Income/NoPeople)

scores <- quantile(table1$AverageIncome)
q1 <- quantile(table1$AverageIncome, 0.25) 
q2 <- quantile(table1$AverageIncome, 0.50)
q3 <- quantile(table1$AverageIncome, 0.75)
q4 <- quantile(table1$AverageIncome, 1)

table1 <- table1%>% mutate(Q25 = ifelse(AverageIncome <= q1,1,0), Q50 = ifelse(AverageIncome > q1 & AverageIncome <= q2,1,0), 
                           Q75 = ifelse(AverageIncome > q2 & AverageIncome <= q3,1,0), Q1 = ifelse(AverageIncome > q3 & AverageIncome <= q4,1,0))

```



# Table 2

## Table 2 Year 2013: Gathering Information for the 2013 Financial Year
```{r,  warning = FALSE}
urldss0 <- "https://data.gov.au/dataset/cff2ae8a-55e4-47db-a66d-e177fe0ac6a0/resource/6b2343f4-09d3-4b03-a1fe-41897f710c6c/download/dssdemographicsdecember2013.xlsx"
p7 <- tempfile()

GET(urldss0, write_disk(tf7 <- tempfile(fileext = ".xlsx")))

dss2013 <- as.data.frame(read_excel(tf7, sheet = 7, col_names = TRUE, skip = 2))

d2013 <- rep("2013", length(dss2013$Postcode))

dss2013 <- dss2013%>%mutate(Year = 2013)

names2013 <- colnames(dss2013)

```

## Table 2 Year 2014: Gathering Information for the 2014 Financial Year
```{r,  warning = FALSE}
urldss1 <- "https://data.gov.au/dataset/cff2ae8a-55e4-47db-a66d-e177fe0ac6a0/resource/f0615bb3-463f-4352-902c-0b6bb0e22e7d/download/june-2014-dss-demographics-suppressed-2014-final-ftb.xlsx"
p8 <- tempfile()

GET(urldss1, write_disk(tf8 <- tempfile(fileext = ".xlsx")))

dss2014 <- as.data.frame(read_excel(tf8, sheet = 7, col_names = TRUE, skip = 2))

d2014 <- rep("2014", length(dss2014$Postcode))

dss2014 <- dss2014%>%mutate(Year = 2014)

colnames(dss2014) <- names2013

```

## Table 2 Year 2015: Gathering Information for the 2015 Financial Year
```{r,  warning = FALSE}
urldss2 <- "https://data.gov.au/dataset/cff2ae8a-55e4-47db-a66d-e177fe0ac6a0/resource/e6457899-378e-406f-8027-a6ee8a19eec6/download/dss-demographics-june-2015.xlsx"
p9 <- tempfile()

GET(urldss2, write_disk(tf9 <- tempfile(fileext = ".xlsx")))

dss2015 <- as.data.frame(read_excel(tf9, sheet = 9, col_names = TRUE, skip = 2))

d2015 <- rep("2015", length(dss2015$Postcode))

dss2015 <- dss2015%>%mutate(Year = 2015)

colnames(dss2015) <- names2013

str(dss2015)


```

## Table 2 Year 2016: Gathering Information for the 2016 Financial Year
```{r,  warning = FALSE}
urldss3 <- "https://data.gov.au/dataset/cff2ae8a-55e4-47db-a66d-e177fe0ac6a0/resource/829fc910-e263-4a59-ad33-1cd1159ce0df/download/dss-demographics-june-2016.xlsx"
p10 <- tempfile()

GET(urldss3, write_disk(tf10 <- tempfile(fileext = ".xlsx")))

dss2016 <- as.data.frame(read_excel(tf10, sheet = 9, col_names = TRUE, skip = 2))

d2016 <- rep("2016", length(dss2016$Postcode))

dss2016 <- dss2016%>%mutate(Year = d2016)

colnames(dss2016) <- names2013


```

## Table 2: Consolidation Payment Demographic Table 2
```{r,  warning = FALSE}
PaymentDemographic<- rbind(dss2013,dss2014, dss2015, dss2016)

PaymentDemographic[PaymentDemographic=="<20"]<-10

PaymentDemographic[is.na(PaymentDemographic)] <- 0

PaymentDemographic[,2:29] <- sapply(PaymentDemographic[, 2:29], as.numeric)

table2 <- PaymentDemographic%>%mutate(Postcode = as.character(Postcode))


```


## Merging Table1 and Table 2

```{r,  warning = FALSE}

mergetables <- merge(x= table1, y = table2, by.x = c('Postcode','Year'), by.y = c('Postcode','Year'), all.x = TRUE)

mergetables[is.na(mergetables)] <- 0

write.csv(mergetables, "data_consolidation.csv")

```


